---
title: "Regression Models Project"
author: "Eric VACHON"
date: "June 2015"
output: pdf_document
geometry: margin = 1.5cm
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data(mtcars)
```

## Summary
In this project, we work for a magazin about automobile industry : Motor Trend. They ask us two questions :
- Question 1 : Is an automatic or manual transmission better for MPG ?
- Question 2 : Quantify the MPG difference between automatic and manual transmissions ?

And to answer this two questions, we are going to use the R dataset "mtcars", data which was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

## Dataset : mtcars
In first look at the dataset : data frame with 32 observations on 11 variables : **mpg** (Miles/(US) gallon), **cyl** (Number of cylinders), **disp** (Displacement), **hp** (Gross horsepower), **drat** (Rear axle ratio), **wt** (Weigh), **qsec** (1/4 mile time), **vs** (V/S), **am** (Transmission (0 = automatic, 1 = manual)), **gear** (Number of forward 
gears) and **carb** (Number of carburetors).
```{r, echo=FALSE}
head(mtcars,3)
```

## Question 1 : Is an automatic or manual transmission better for MPG ?
First we must analyse the data for automatic and manual transmmission 
(cf plot 1 in Appendix) and run a t.test between this 2 subsets :
```{r, echo=FALSE}
t.test(mtcars[mtcars$am==0,"mpg"],mtcars[mtcars$am==1,"mpg"],conf.level=0.95)
```
The p value is less than 0.05, 0 is not include in the 95 percent confidence interval, the mean of automatic transmission (`r round(mean(subset(mtcars,am==0,mpg)$mpg),2)`) is less than the manual (`r round(mean(subset(mtcars,am==1,mpg)$mpg),2)`), and (plot 1) we can see that automatic transmission consume globally less than the manual.

So we can conclude that : **automatic transmission is better for MPG**.

## Question 2 : Quantify the MPG difference between automatic and manual transmissions ?
In this section we are going to use in first a simple model of regression : **Is type of transmission can explain the mpg ?**
```{r}
simpleLM<-lm(mpg~am, data = mtcars)
summary(simpleLM)$r.square
summary(simpleLM)$coefficients
confint(simpleLM)
```
So R-squared is low (`r summary(simpleLM)$r.square`), p value is less than 0.05, 0 is not include in the confidence interval : So `r round(summary(simpleLM)$r.square*100,2)`% of the variation of mpg can be explain by the type of transmission ... **we can find a better model**

So we are going to find a another variables to quantify better the MPG difference between automatic and manual transmissions.

First look at the correlation between all the 11 variables (cf plot 3 in the appendix), and we must find the better combinaison of variables to fin the best model.  To do this we are going to use the R-step function  :

```{r}
multipleLM <- step(lm(data = mtcars, mpg ~ .),trace=0)
summary(multipleLM)$r.square
summary(multipleLM)$coefficients
confint(multipleLM)
```
Now we have a model that `r round(summary(multipleLM)$r.square*100,2)`% of the variation of mpg can be explain by : the weight + 1/4 mile time + type of transmission.  
Now look at the normality of residuals (plot 4) : the residuals seem to be normal (an anova help too).  
In fact this model explain that :   
1. We estimate an expected 1.22 miles/gallon increase in fuel consumption for every 1 seconde increase in speed to do 1/4 miles in holding the remaining variables constan t(wt and am) : plot 3.  
2. We estimate an expected 3.91 miles/gallon decrease in fuel consumption for every lb/1000 increase in weight in holding the remaining variables constan t(qsec and am) : plot 4.  
3. We estimlate in mean an expected increase by 2.9 mpg the fuel consumption if we use a manual transmission instead of an automatic transmission (plot 3 and 4).  

```{r}
```
## Appendix
Libraries
```{r,message=FALSE, warning=FALSE}
library(ggplot2);library(corrplot);library(plyr)
mtcars$TypeTransmission <- factor(mtcars$am)
mtcars$TypeTransmission <- revalue(mtcars$TypeTransmission, c("0"="automatic", "1"="manual"))
```
```{r,message=FALSE, warning=FALSE , fig.width=6, fig.height=4}
ggplot(mtcars, aes(x=TypeTransmission, y=mpg, fill=TypeTransmission)) + geom_boxplot() +
       labs(x="Type of transmission",y = "MPG",title = "Plot 1 : MPG by type of transmission")+
       guides(fill=guide_legend(title="Type of transmission"))
```
```{r,message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
par(cex = 0.7)
corrplot.mixed(cor(mtcars[,c(1:11)]),order = "FPC",mar=c(0,0,1,0),
               title="Plot 2 : Correlation between variables")
```
```{r,message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
ggplot(mtcars, aes(x = qsec, y = mpg, group = 
  TypeTransmission))+geom_smooth(method = "lm")+
       geom_point(size = 7, aes(colour = TypeTransmission), alpha=0.5) +
       labs(x = "time to 1/4 miles",y = "MPG", title = "Plot 3 : MPG by time to 1/4 miles and type of transmission")
```
```{r,message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
ggplot(mtcars, aes(x = wt, y = mpg, group=TypeTransmission))+geom_smooth(method = "lm")+
       geom_point(size = 7, aes(colour = TypeTransmission), alpha=0.5) +
       labs(x = "Weight (lb/1000)",y = "MPG", title = "Plot 4 : MPG by weigth and type of transmission")
```
```{r,message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
qqnorm(resid(multipleLM), col = 'blue',main = 'Plot 4 : Distribution of the residuals')
qqline(resid(multipleLM), col = 'red')  
```   
  
## End of the document